---
title: "Plotting Spot Rate Curves"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting Spot Rate Curves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

Visualization of yield curves is fairly relevant and the package fixedincome brings the usual
`plot` function to create plots with the SpotRateCurve objects and also a ggplot2 set of methods
that helps with the visualization of the term structures using all ggplot2 features.

Before the start let's get the necessary packages loaded.

```{r message=FALSE, warning=FALSE}
library(rb3)
library(dplyr)
library(fixedincome)
library(ggplot2)
```

As the second step, the yield curve must be created and this is done with data from {rb3} package.

```{r}
refdate <- as.Date("2022-08-09")
yc_ <- yc_get(refdate)
fut_ <- futures_get(refdate)
yc_ss <- yc_superset(yc_, fut_)
yc <- bind_rows(
  yc_ss |> slice(1),
  yc_ss |> filter(!is.na(symbol))
) |>
  filter(!duplicated(biz_days))

curve <- spotratecurve(
  yc$r_252, yc$biz_days, "discrete", "business/252", "Brazil/ANBIMA",
  refdate = refdate
)
```

### `base::plot`
```{r fig.width=10, fig.height=5}
plot(curve)
```

```{r fig.width=10, fig.height=5}
plot(curve, show_forward = TRUE)
```

```{r fig.width=10, fig.height=5}
curve |>
  fixedincome::first("3 years") |>
  plot(curve, show_forward = TRUE)
```

```{r fig.width=10, fig.height=5}
curve_2y <- curve |> fixedincome::first("2 years")
interpolation(curve_2y) <- interp_naturalspline()
plot(curve_2y, use_interpolation = TRUE)
```

```{r fig.width=10, fig.height=5}
plot(curve_2y, use_interpolation = TRUE, show_forward = TRUE)
```

```{r fig.width=10, fig.height=5}
plot(curve_2y, use_interpolation = TRUE, show_forward = TRUE, legend_location = "bottomleft")
```

### ggplot2

```{r fig.width=10, fig.height=5}
ggspotratecurveplot(curve,
  title = "DI1 spot rates", subtitle = format(refdate), caption = "Data from {rb3} package"
)
```

```{r fig.width=10, fig.height=5}
ggspotratecurveplot(curve,
  title = "DI1 spot rates", subtitle = format(refdate), caption = "Data from {rb3} package",
  curve.x.axis = "terms"
)
```

```{r fig.width=10, fig.height=5}
ggspotratecurveplot(curve,
  title = "DI1 spot rates", subtitle = format(refdate), caption = "Data from {rb3} package"
) +
  autolayer(forwardrate(curve), size = 1)
```

```{r fig.width=10, fig.height=5}
ggspotratecurveplot(curve_2y,
  title = "DI1 spot rates", subtitle = format(refdate), caption = "Data from {rb3} package",
  curve.interpolation = TRUE
)
```

```{r fig.width=10, fig.height=5}
autoplot(curve_2y, curve.geom = "point") +
  autolayer(curve_2y, curve.geom = "line", curve.name = "Interpolation", curve.interpolation = TRUE)
```

```{r fig.width=10, fig.height=5}
terms_ <- do.call(seq, as.list(range(curve_2y@terms)))
curve_2y_interp <- curve_2y[[terms_]]
autoplot(curve_2y, curve.geom = "point") +
  autolayer(curve_2y_interp, curve.geom = "line") +
  autolayer(forwardrate(curve_2y_interp),
    curve.geom = "line", curve.name = "Forward Rate"
  )
```
